---
title: "ESM 244 Lab 1"
author: "Laurel Wee"
date: "1/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(sf)
library(tmap)
```

### Read in SF trees data

```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))
```

### Basic wrangling reminders

Refresh some skills for data wrangling & summary statistics using functions in the 'dplyr' package. 

Find the top 5 highest observations of tress by legal_status, do some wrangling, make a graph.

```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>% 
  slice_max(tree_count, n=5)
```

Make a graph of those top 5 observations by legal status.
```{r}
ggplot(data = top_5_status, aes(x= fct_reorder(legal_status, tree_count), y= tree_count))+
  geom_col()+
  labs(x= "Legal Status", y= "Tree Count")+
  coord_flip()+
  theme_minimal()
```

### More wrangling examples

Only want to keep observations (rows) for Blackwood Acacia trees.

```{r}
blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, latitude, longitude)

ggplot(data = blackwood_acacia, aes(x = longitude, y = latitude)) +
  geom_point()
```






